---
title: "EDA"
output: html_document
date: '2022-05-27'
---

```{r setup}
library(ggplot2)
library(psych)
library(plyr)

setwd('C:/Users/Noel/Desktop/UM - Master of Data Science/WQD7001 - Principles of Data Science/Group Project')
df<-read.csv(file = 'student-mat.csv')
# df<-read.csv(file = "C:/Users/Noel/Desktop/UM - Master of Data Science/WQD7001 - Principles of 
#                      Data Science/Group Project/student-mat.csv")

# add a new column with the mean of G1 G2 and G3
df$score_mean <- rowMeans(subset(df, select = c(G1, G2, G2)), na.rm = TRUE)
# df = subset(df, select = -c(mean)) #drop mean column
describe(df$score_mean)
df
```

################################################################################
Q1 Do parents’ education background correlate with student’s grade in the Math subject?
Ans: no relation between edu vs grade
     but
     failure, studytime, higher shows positive correlation vs grade 

```{r}
# avg score based on number of distinct in Medu/student 
Medu_score = c()
for (x in 0:4)
{
  Medu_score_mean = df[df$Medu==x, c("Medu" , "score_mean")]
  Medu_score = append(Medu_score,mean(Medu_score_mean$score_mean))
  cat(x, mean(Medu_score_mean$score_mean),"\n")
}
barplot(Medu_score,names.arg=c("0","1","2","3","4"),main = "avg test scores vs Medu",
xlab = "Medu",
ylab = "avg test scores",
col = c("red", "green", "yellow", "blue", "purple"))
```

```{r}
# avg score based on number of distinct in Fedu/student 
Fedu_score = c()
for (x in 0:4)
{
  Fedu_score_mean = df[df$Fedu==x, c("Fedu" , "score_mean")]
  Fedu_score = append(Fedu_score, mean(Fedu_score_mean$score_mean))
  cat(x, mean(Fedu_score_mean$score_mean),"\n")
}
barplot(Fedu_score,names.arg=c("0","1","2","3","4"),main = "avg test scores vs Fedu",
xlab = "Fedu",
ylab = "avg test scores",
col = c("red", "green", "yellow", "blue", "purple"))
```

```{r}
# failures vs grade
ggplot(df, aes(x=score_mean, y=failures, color=failures)) + geom_point(size=3)
```

```{r}
# studytime vs grade
ggplot(df, aes(x=score_mean, y=studytime, color=studytime)) + geom_point(size = 3)
```

```{r}
# grade vs higher
higher_yes = df[df$higher == 'yes', c("higher" , "score_mean")]
# count(higher_yes$higher == 'yes')
yes=mean(higher_yes$score_mean)
higher_no = df[df$higher == 'no', c("higher" , "score_mean")]
# count(higher_no$higher == 'no')
no=mean(higher_no$score_mean)
cat("yes",yes,"no",no)
higheryesno = data.frame(higher = c('yes','no'), score = c(yes,no))
# higheryesno
ggplot(higheryesno, aes(higher,score,fill = higher)) + geom_bar(stat = "identity") + geom_text(aes(label = signif(score)), nudge_y = 1)
```

################################################################################
Q2 Do parents’ education background correlate with student’s past class failures?
Ans: Negative correlation between edu vs failure

```{r}
#avg of number of failures vs Medu
count(df$failures)
Medu_failure = c()
for (x in 0:4)
{
  Medu_failure_mean = df[df$Medu==x, c("Medu" , "failures")]
  Medu_failure = append(Medu_failure, mean(Medu_failure_mean$failures))
  cat(x, mean(Medu_failure_mean$failures),"\n")
}
barplot(Medu_failure,names.arg=c("0","1","2","3","4"),main = "avg number of failures vs Medu",
xlab = "Medu",
ylab = "avg number of failures",
col = c("red", "green", "yellow", "blue", "purple"))
```

```{r}
#avg of number of failures vs Fedu
Fedu_failure = c()
for (x in 0:4)
{
  Fedu_failure_mean = df[df$Fedu==x, c("Fedu" , "failures")]
  Fedu_failure = append(Fedu_failure, mean(Fedu_failure_mean$failures))
  cat(x, mean(Fedu_failure_mean$failures),"\n")
}
barplot(Fedu_failure,names.arg=c("0","1","2","3","4"),main = "avg number of failures vs Fedu",
xlab = "Fedu",
ylab = "avg number of failures",
col = c("red", "green", "yellow", "blue", "purple"))
```

################################################################################
Q3 Do parents’ education background affect student’s total study time?
Ans: Positive correlation between edu vs studytime 

```{r}
#total study time vs Medu
Medu_studytime = c()
for (x in 0:4)
{
  Medu_studytime_sum = df[df$Medu==x, c("Medu" , "studytime")]
  Medu_studytime = append(Medu_studytime, sum(Medu_studytime_sum$studytime))
  cat(x, sum(Medu_studytime_sum$studytime),"\n")
}
barplot(Medu_studytime,names.arg=c("0","1","2","3","4"),main = "study time vs Medu",
xlab = "Medu",
ylab = "total study time",
col = c("red", "green", "yellow", "blue", "purple"))
```


```{r}
#total study time vs Fedu
Fedu_studytime = c()
for (x in 0:4)
{
  Fedu_studytime_sum = df[df$Fedu==x, c("Fedu" , "studytime")]
  Fedu_studytime = append(Fedu_studytime, sum(Fedu_studytime_sum$studytime))
  cat(x, sum(Fedu_studytime_sum$studytime),"\n")
}
barplot(Fedu_studytime,names.arg=c("0","1","2","3","4"),main = "study time vs Fedu",
xlab = "Fedu",
ylab = "total study time",
col = c("red", "green", "yellow", "blue", "purple"))
```

################################################################################
Q4 Do students with highly educated parents tend to be more ambitious and pursue higher education?
Ans: Positive correlation between edu vs higher

```{r}
#total higher vs Medu
Medu_higher = c()
for (x in 0:4)
{
  Medu_higher_count = df[df$Medu==x, c("Medu" , "higher")]
  count_higher = count(Medu_higher_count$higher == 'yes')
  count_higher = data.frame(count_higher)
  Medu_higher = append(Medu_higher, count_higher[count_higher$x=='TRUE', c("freq")])
  cat(x, count_higher[count_higher$x=='TRUE', c("freq")],"\n")
}
barplot(Medu_higher,names.arg=c("0","1","2","3","4"),main = "total higher vs Medu",
xlab = "Medu",
ylab = "total higher",
col = c("red", "green", "yellow", "blue", "purple"))
```


```{r}
#total higher vs Fedu
Fedu_higher = c()
for (x in 0:4)
{
  Fedu_higher_count = df[df$Fedu==x, c("Fedu" , "higher")]
  count_higher = count(Fedu_higher_count$higher == 'yes')
  count_higher = data.frame(count_higher)
  Fedu_higher = append(Fedu_higher, count_higher[count_higher$x=='TRUE', c("freq")])
  cat(x, count_higher[count_higher$x=='TRUE', c("freq")],"\n")
}
barplot(Fedu_higher,names.arg=c("0","1","2","3","4"),main = "total higher vs Fedu",
xlab = "Fedu",
ylab = "total higher",
col = c("red", "green", "yellow", "blue", "purple"))
```



```{r}
# grade vs paid
paid_yes = df[df$paid == 'yes', c("paid" , "score_mean")]
# count(paid_yes$paid == 'yes')
yes=mean(paid_yes$score_mean)
paid_no = df[df$paid == 'no', c("paid" , "score_mean")]
# count(paid_no$paid == 'no')
no=mean(paid_no$score_mean)
cat("yes",yes,"no",no)
paidyesno = data.frame(boolean = c('yes','no'), score = c(yes,no))
ggplot(paidyesno, aes(boolean,score)) + geom_bar(stat = "identity") + geom_text(aes(label = signif(score)), nudge_y = 1)
```
